---
title: "Process error from F-forward simulations"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Process error from F-forward simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5
)
```

Worked example: build a life-history OM, project it with `ffwd()` under known
\(F/F_{\mathrm{MSY}}\) trajectories, recover process-error residuals with
`pe()`, and fit JABBA with `proc.dyn`.

# Setup

```{r setup}
library(FLCore)
library(FLBRP)
library(FLife)
library(FLasher)
library(FLRebuild)
library(ggplot2)
set.seed(1)
nits <- 8
it   <- 1
```

# Life-history OM

```{r om}
par <- lhPar(FLPar(linf = 80))
eq  <- lhEql(par)
refpts(eq)[c("msy", "crash"), c("harvest", "ssb", "yield")]
```

# Constant \(F/F_{\mathrm{MSY}}\)

```{r constant-f}
stk <- eq
FLBRP::fbar(stk)[] <- 1
stk <- as(stk, "FLStock")

mlt <- seq(0.01, 2, length.out = 11)
stk <- propagate(stk, length(mlt))
ftar <- (FLCore::fbar(eq) %=% 1) %*%
  as.FLQuant(mlt, dimnames = list(iter = seq_along(mlt))) *
  refpts(eq)["msy", "harvest"]
devs <- rlnorm(1, rec(stk)[, , , , , 1] %=% 0, 0.3)

stk.f <- window(ffwd(stk, fbar = ftar[, -1], sr = eq, deviances = devs),
                start = 75)
pe.f  <- pe(stk.f, eq, location = "after", scale = "log")

pe_vs_f <- data.frame(
  F  = mlt,
  PE = c(apply(pe.f, 6, function(x) stats::sd(x, na.rm = TRUE)))
)

ggplot(pe_vs_f, aes(F, PE)) +
  geom_line() +
  geom_point() +
  labs(x = expression(F / F[MSY]), y = "SD(log PE)") +
  theme_bw()
```

PE scale rises with \(F\).

# Increasing vs declining \(F\)

```{r trajectories}
fwd_pe <- function(f_mult, nits, rec_sd = 0.3) {
  par <- lhPar(FLPar(linf = 80))
  eq  <- lhEql(par)
  stk <- eq
  ftar <- FLQuant(f_mult) * refpts(eq)["msy", "harvest"]
  fbar(stk) <- ftar
  stk <- propagate(as(stk, "FLStock"), nits)
  devs <- rlnorm(nits, rec(stk) %=% 0, rec_sd)
  stk <- ffwd(stk, fbar = fbar(stk)[, -1], sr = eq, deviances = devs)
  list(stk = stk, eq = eq,
       pe = pe(stk, eq, location = "after", scale = "log"))
}

inc <- fwd_pe(c(rep(0.01, 30), seq(0.01, 2, length.out = 41)), nits)
dec <- fwd_pe(c(rep(2, 30),    seq(2, 0.01, length.out = 41)), nits)
```

```{r pe-sd-year}
sd_by_year <- function(x, start = 25) {
  y <- apply(window(x, start = start), 2, function(z) stats::sd(z, na.rm = TRUE))
  df <- as.data.frame(y)
  names(df)[names(df) == "data"] <- "PE"
  df
}

sd_df <- rbind(
  transform(sd_by_year(inc$pe), trajectory = "Increase"),
  transform(sd_by_year(dec$pe), trajectory = "Decline")
)

ggplot(sd_df, aes(year, PE, colour = trajectory)) +
  geom_line() +
  labs(x = "Year", y = "SD(log PE)") +
  theme_bw()
```

# `plotPe()`

```{r plotpe}
ts1 <- FLQuants(
  pe    = iter(inc$pe, it),
  bbmsy = iter(ssb(inc$stk) %/% refpts(inc$eq)["msy", "ssb"], it)
)
plotPe(ts1)
```

# Fit JABBA

One realisation (`it`) of the increasing-$F$ stock.

```{r jabba, eval = requireNamespace("JABBA", quietly = TRUE)}
stk.j <- iter(inc$stk, it)
fit <- runJABBA(stk.j, inc$eq, index = ebiomass, quick = TRUE, nc = 2)
if (!is.null(fit) && !is.null(fit$timeseries)) {
  plotPe(jabbaTs(fit))
  plotJabbaTs(fit)
  plotJabbaTs(fit, qname = "bbmsy",
              om = FLQuants(
                `OM SSB` = ssb(stk.j) %/% refpts(inc$eq)["msy", "ssb"],
                `OM EB`  = ebiomass(stk.j) %/% c(refptsEB(inc$eq)["msy", "eb"])))
  jabba2biodyn(fit)
}
```

See `vignettes/jabba-workflow.Rmd` for priors and `jabba2biodyn()`.
