---
title: "JABBA Workflow in FLRebuild"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{JABBA Workflow in FLRebuild}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Purpose

This vignette shows the recommended JABBA workflow now implemented in `FLRebuild`:

1. Build model inputs with `jabbaInput()`
2. Build priors with `jabbaPriors()`
3. Fit with `runJABBA()`

It also shows where to get the wider stock reference table (`icesdata::refs()`), which is separate from compact JABBA priors.

# Prerequisites

```{r setup, eval=FALSE}
library(FLRebuild)
library(icesdata)
```

`runJABBA()` relies on the `JABBA` package and an external model toolchain. If unavailable, you can still prepare and inspect inputs/priors.

# Example data

```{r data}
library(icesdata)
data(icesdata)

stk <- icesdata[[1]]
```

# Step 1: Build JABBA input

`jabbaInput()` converts an `FLStock` into a list with:

- `catch`: data frame with `year`, `catch`
- `index`: data frame with `year`, `index`

```{r input}
inp <- FLRebuild::jabbaInput(stk, index = "ebiomass")

str(inp, max.level = 1)
head(inp$catch)
head(inp$index)
```

You can pass `index` as:

- a function name (character), e.g. `"ebiomass"`
- a function taking `object` and returning an index
- an `FLQuant`
- a data frame with columns `year`, `index`

# Step 2: Build priors

`jabbaPriors()` returns the compact priors object expected by JABBA:

- `pr` (`FLPar`): prior means
- `pr.sd` (`FLPar`): prior standard deviations

Available methods:

- `"ices"`
- `"fishlife"`
- `"flbrp"`

```{r priors}
priIces <- FLRebuild::jabbaPriors(stk, method = "ices")
priFishlife <- FLRebuild::jabbaPriors(stk, method = "fishlife")

priIces$pr
priIces$pr.sd
```

## About `method = "flbrp"`

```{r priors-flbrp, eval=FALSE}
priFlbrp <- FLRebuild::jabbaPriors(stk, method = "flbrp")
```

The `flbrp` route uses `eql()`/`brp()` output and extracts reference points required for JABBA priors.

# Step 3: Run JABBA

Run directly from `FLStock`:

```{r run-stock, eval=FALSE}
fit <- FLRebuild::runJABBA(
  stk,
  method = "ices",
  quick = TRUE,
  nc = 3
)
```

Or run from prepared data frames and explicit priors:

```{r run-df, eval=FALSE}
fitDf <- FLRebuild::runJABBA(
  object = inp$catch,
  priors = priIces,
  index  = inp$index,
  quick  = TRUE,
  nc     = 3
)
```

# Post-fit helpers

`FLRebuild` provides convenience tools for JABBA outputs:

```{r postfit, eval=FALSE}
ext <- FLRebuild::jabbaExtract(fit)

names(ext)
# "posteriors" "trajectory" "priors"

# Optional diagnostics/plots:
# FLRebuild::jabbaDiagnostics(fit)
# FLRebuild::jbplot(list(base = fit))
```

# Wide reference table is now separate

The old wide table (benchmark/eqsim/fishlife + depletion summaries) is not `jabbaPriors()`.
Use `icesdata::refs()`:

```{r refs-table}
rtab <- icesdata::refs(icesdata)

head(rtab[, c(".id", "r", "shape", "psi",
              "ssb.minyear", "ssb.maxyear",
              "ssb.minyear.year", "ssb.maxyear.year")])
```

# Recommended pattern

Use:

- `FLRebuild::jabbaPriors()` for model-ready priors (`pr`, `pr.sd`)
- `icesdata::refs()` for broad stock reference summaries

This keeps model fitting inputs strict and predictable while preserving the richer reference table workflow.
