---
title: "JABBA Process-Error Diagnostics"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{JABBA Process-Error Diagnostics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Purpose

Extract the JABBA process-error time series into `FLQuants` and produce the
standard four-panel residual diagnostics used for process-error review.

# Setup

```{r setup}
library(FLCore)
library(FLRebuild)
library(ggplot2)
```

# From a JABBA fit

After fitting with `runJABBA()` / `jabba()`, process deviations are in
`fit$timeseries[,"mu","procB"]`. Use `proc.dyn = TRUE` to scale process SD
by \(B/B_{\mathrm{MSY}}\) (\(\sigma_t = \sigma \times \mathrm{relsig}(B/B_{\mathrm{MSY}})\)).

```{r jabba-ts, eval=FALSE}
fit   <- runJABBA(stk, eq, index = ebiomass, quick = TRUE)
fitPE <- runJABBA(stk, eq, index = ebiomass, proc.dyn = TRUE, quick = TRUE)

jbTs   <- jabbaTs(fit)
jbTsPE <- jabbaTs(fitPE)
names(jbTs)
# stock harvest bbmsy ffmsy pe sprod

plotPe(jbTs)
plotPe(jbTsPE)

plotJabbaTs(fit)
plotJabbaTs(list(JABBA = fit, `JABBA-PE` = fitPE), qname = "bbmsy")
```

`jabbaTs()` maps medians. `jabbaTsCI()` / `plotJabbaTs()` add JABBA's 95% credible intervals (`lci`/`uci`):

| FLQuants name | JABBA quantity |
|---------------|----------------|
| `stock`       | `B`            |
| `harvest`     | `F`            |
| `bbmsy`       | `BBmsy`        |
| `ffmsy`       | `FFmsy`        |
| `pe`          | `procB`        |
| `sprod`       | `SPt`          |

# Four-panel diagnostics

```{r plot-pe-fit, eval=FALSE}
plotPe(jbTs)
# or directly:
# plotPe(fit)
```

Panels:

1. **Time** — residuals vs year, with optional STARS regime polygons (`rod()`).
2. **Status** — residuals vs \(B/B_{MSY}\) (bias / heteroscedasticity).
3. **ACF** — temporal dependence beyond white noise.
4. **Histogram** — residual distribution vs a normal density.

# Illustration without a JABBA run

```{r simulate}
set.seed(1)
yrs <- 1980:2020
n <- length(yrs)
pe <- rnorm(n, 0, 0.15)
pe[25:35] <- pe[25:35] + 0.35
bbmsy <- pmax(0.2, 1.2 - seq_len(n) / n + rnorm(n, 0, 0.05))

jbTs <- FLQuants(
  pe    = as.FLQuant(data.frame(year = yrs, data = pe)),
  bbmsy = as.FLQuant(data.frame(year = yrs, data = bbmsy))
)

plotPe(jbTs)
```

# Relation to `jabbaPE()`

- `jabbaTs()` / `plotPe()` use JABBA's estimated state process deviations (`procB`).
- `jabbaPE()` builds the production curve and derives residuals from
  \(B_{t+1}\) vs expected \(B_t + P(B_t) - C_t\).

Use `jabbaTs` for diagnostics of the fitted state-space process; use `jabbaPE`
when you want production-function phase plots and derived surplus-production
residuals.

# See also

- `vignettes/jabba-workflow.Rmd` for fitting
- `vignettes/ffwd-process-error.Rmd` for the F-forward OM worked example
- `rod()` / `rodFn()` for STARS regime detection
- `jabbaPE()` for production-function residuals
