---
title: "JABBA Workflow in FLRebuild"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{JABBA Workflow in FLRebuild}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Purpose

This vignette shows the recommended JABBA workflow now implemented in `FLRebuild`:

1. Build model inputs with `jabbaInput()`
2. Build priors with `jabbaPriors()`
3. Fit with `runJABBA()`

It also shows where to get the wider stock reference table (`icesdata::refs()`), which is separate from compact JABBA priors.

# Prerequisites

```{r setup, eval=FALSE}
library(FLRebuild)
library(icesdata)
```

`runJABBA()` relies on the `JABBA` package and an external model toolchain. If unavailable, you can still prepare and inspect inputs/priors.

# Example data

```{r data}
library(icesdata)
data(icesdata)

stk <- icesdata[[1]]
# eq <- eql(stk)   # or an FLBRP you already have (jabba-f)
```

# Step 1: Build JABBA input

`jabbaInput()` converts an `FLStock` into a list with:

- `catch`: data frame with `year`, `catch`
- `index`: data frame with `year`, `index`

```{r input}
inp <- FLRebuild::jabbaInput(stk, index = "ebiomass")

str(inp, max.level = 1)
head(inp$catch)
head(inp$index)
```

You can pass `index` as:

- a function name (character), e.g. `"ebiomass"`
- a function taking `object` and returning an index
- an `FLQuant`
- a data frame with columns `year`, `index`

# Step 2: Build priors

`jabbaPriors()` returns the compact priors object expected by JABBA:

- `pr` (`FLPar`): prior means
- `pr.sd` (`FLPar`): prior standard deviations

Available methods:

- `"ices"`
- `"fishlife"`
- `"flbrp"`

```{r priors}
priIces <- FLRebuild::jabbaPriors(stk, method = "ices")
priFishlife <- FLRebuild::jabbaPriors(stk, method = "fishlife")

priIces$pr
priIces$pr.sd
```

## About `method = "flbrp"`

```{r priors-flbrp, eval=FALSE}
priFlbrp <- FLRebuild::jabbaPriors(stk, method = "flbrp")
```

The `flbrp` route uses `eql()`/`brp()` output and extracts reference points required for JABBA priors.

# Step 3: Run JABBA

From an \code{FLStock} and an \code{FLBRP}: catch and the biomass index come
from the stock; \(F_{\mathrm{MSY}}\), \(B_{\mathrm{MSY}}\) and \(B_0\) come
from the equilibrium object. The default index is exploitable biomass.

```{r run-stock-eq, eval=FALSE}
fit  <- FLRebuild::runJABBA(stk, eq, index = ebiomass)
fitF <- FLRebuild::runJABBA(stk, eq, index = ebiomass, f = ffmsy)
bd   <- FLRebuild::runJABBA(stk, eq, index = ebiomass, output = "mpb")
bd   <- FLRebuild::jabba2biodyn(fit)   # same, from an existing fit

# Process error: estimate SD (default), fix it, or scale by B/BMSY
fitFix <- FLRebuild::runJABBA(stk, eq, index = ebiomass, sigma.proc = 0.07)
fitPE  <- FLRebuild::runJABBA(stk, eq, index = ebiomass, proc.dyn = TRUE)
bdPE   <- FLRebuild::jabba2biodyn(fitPE)
```

`f = "ffmsy"` is the jabba-f auxiliary series \(F/F_{\mathrm{MSY}}\) from
`fbar` and `eq`. Absolute \(F\) is `f = fbar` (or `f = "f"`).

Process error (`sigma.proc = TRUE` by default) is estimated with inverse-gamma
prior `igamma = c(3, 0.1)`. Pass a number to fix it. `proc.dyn = TRUE` scales
that SD by depletion: \(\sigma_t = \sigma \times \mathrm{relsig}(B/B_{\mathrm{MSY}})\)
(GitHub JABBA with `proc.dyn`). The curve coefficients default to JABBA's
ICES values; override with `pe = c(a = 0.2, c = 0.35, d = 2)`.

```{r run-pe-args, eval=FALSE}
fitPE <- FLRebuild::runJABBA(stk, eq, index = ebiomass, proc.dyn = TRUE,
                              pe = c(a = 0.2, c = 0.35, d = 2))
```

From an \code{FLStock} alone, using a prior method:

```{r run-stock, eval=FALSE}
fit <- FLRebuild::runJABBA(
  stk,
  method = "ices",
  quick = TRUE,
  nc = 3
)
```

Or run from prepared data frames and explicit priors:

```{r run-df, eval=FALSE}
fitDf <- FLRebuild::runJABBA(
  object = inp$catch,
  priors = priIces,
  index  = inp$index,
  quick  = TRUE,
  nc     = 3
)
```

# Post-fit helpers

`FLRebuild` provides convenience tools for JABBA outputs:

```{r postfit, eval=FALSE}
ext <- FLRebuild::jabbaExtract(fit)

names(ext)
# "posteriors" "trajectory" "priors"

# Optional diagnostics/plots:
# FLRebuild::jabbaDiagnostics(fit)
# FLRebuild::jbplot(list(base = fit))

# Process-error time series and four-panel diagnostics:
# ts <- FLRebuild::jabbaTs(fit)
# FLRebuild::plotPe(ts)
# FLRebuild::plotJabbaTs(fit)


# Constant vs depletion-dependent process error

A single fit with depletion-dependent PE:

```{r run-pe, eval=FALSE}
fitPE <- FLRebuild::runJABBA(stk, eq, index = ebiomass, proc.dyn = TRUE)
fitPE <- FLRebuild::runJABBA(stk, eq, index = ebiomass, proc.dyn = TRUE,
                              pe = c(a = 0.2, c = 0.35, d = 2))
tsPE  <- FLRebuild::jabbaTs(fitPE)
plotPe(tsPE)
bdPE  <- FLRebuild::jabba2biodyn(fitPE)
```

`compareJabbaPE()` fits constant PE and `proc.dyn = TRUE` on the same stock
and priors:

```{r compare-pe, eval=FALSE}
cmp <- FLRebuild::compareJabbaPE(stk, eq = eq, quick = TRUE, nc = 3)
cmp$status
plot(cmp)
plotPe(cmp$ts$constant)
plotPe(cmp$ts$proc.dyn)
```

The relative-SD curve is `relsig()` (ICES coefficients hardcoded in JABBA).

# Wide reference table is now separate

The old wide table (benchmark/eqsim/fishlife + depletion summaries) is not `jabbaPriors()`.
Use `icesdata::refs()`:

```{r refs-table}
rtab <- icesdata::refs(icesdata)

head(rtab[, c(".id", "r", "shape", "psi",
              "ssb.minyear", "ssb.maxyear",
              "ssb.minyear.year", "ssb.maxyear.year")])
```

# Recommended pattern

Use:

- `FLRebuild::jabbaPriors()` for model-ready priors (`pr`, `pr.sd`)
- `icesdata::refs()` for broad stock reference summaries

This keeps model fitting inputs strict and predictable while preserving the richer reference table workflow.
