---
title: "JABBA Priors: Methods and Comparison"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{JABBA Priors: Methods and Comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Purpose

`FLRebuild::jabbaPriors()` provides three ways to construct model-ready JABBA priors from an `FLStock`:

- `"ices"`
- `"fishlife"`
- `"flbrp"`

This vignette shows what each method uses, when to use it, and how to compare outputs.

# Setup

```{r setup}
library(FLRebuild)
library(FLBRP)
library(icesdata)

data(icesdata)
stk <- icesdata[[1]]
```

# Output structure (same for all methods)

Every method returns a list with:

- `pr` (`FLPar`): prior means
- `pr.sd` (`FLPar`): prior standard deviations

Core parameter names expected by JABBA:

- `r`: intrinsic growth
- `psi`: initial depletion
- `shape`: `Bmsy/K`

Optional fields usually included:

- `k`: carrying capacity proxy
- `current`: recent depletion proxy

# Method 1: `method = "ices"`

Uses ICES-style stock attributes:

- `eqsim(object)` for `fmsy`, `bmsy`, `b0`
- stock time series for initial/current depletion summaries

```{r priors-ices}
priIces <- jabbaPriors(stk, method = "ices")
priIces$pr
priIces$pr.sd
```

Use this when your stock has robust ICES reference points available.

# Method 2: `method = "fishlife"`

Uses life-history information:

- `fishlife(object)` for productivity proxy (`r`)
- maximum catch scaling for `k`
- stock time series for depletion summaries

```{r priors-fishlife}
priFishlife <- jabbaPriors(stk, method = "fishlife")
priFishlife$pr
priFishlife$pr.sd
```

Use this when reference-point objects are sparse but life-history attributes are available.

# Method 3: `method = "flbrp"`

Uses the FLBRP route:

- `eql(object, model = sr)` and `brp(...)`
- extracts reference points (`fmsy`, `bmsy`, `b0`) from FLBRP outputs

```{r priors-flbrp}
priFlbrp <- jabbaPriors(stk, method = "flbrp")
priFlbrp$pr
priFlbrp$pr.sd
```

Use this when you want priors tied to the fitted stock-recruit/equilibrium pathway.

# Method 4 (workflow pattern): limited life history -> `lhPar`/`lhEql` -> `method = "flbrp"`

You can also derive `flbrp` priors when stock data are limited by first building
an equilibrium object from life-history inputs.

Conceptually:

1. Build life-history parameters (`lhPar`)
2. Build equilibrium object (`lhEql`)
3. Derive priors via `jabbaPriors(..., method = "flbrp")`

```{r priors-lh-flbrp, eval=FALSE}
# Illustrative pattern (exact arguments depend on your stock/species data):
lhPars <- lhPar(
  # e.g. linf, k, m, mat50, ...
)

lhEq <- lhEql(
  lhPars
  # plus any control options needed by your workflow
)

priFlbrpLh <- jabbaPriors(
  stk,
  method = "flbrp"
  # if your implementation accepts an externally prepared eql/brp object,
  # pass it here; otherwise ensure the stock carries the needed attributes
)
```

Use this pathway when you do not have full assessment/reference-point objects,
but you do have enough life-history information to construct a defensible
equilibrium model.

# Method 5 (workflow pattern): Leslie matrix workflow in `FLife`

Another useful biology-first route is to use a Leslie matrix with `FLife`,
then use those demographic insights to inform priors.

Typical use:

1. Build/obtain an object with enough life-history structure
2. Compute Leslie matrix quantities (e.g. lambda, generation time proxies)
3. Use those results to support prior choices (`r`, `shape`, depletion bounds)
4. Run `jabbaPriors()` with your selected method as the formal prior object

```{r priors-leslie, eval=FALSE}
library(FLife)

# Example pattern using an FLBRP object:
L <- FLife::leslie(rp, fbar = FLCore::FLQuant(c(FLCore::refpts(rp)["msy", "harvest"])))[drop = TRUE]

eig <- eigen(L)$values
lambda <- Re(eig[which.max(Mod(eig))])
rFromLeslie <- log(lambda)

# Use Leslie-derived diagnostics to guide prior expectations/ranges,
# then build compact JABBA priors with the chosen method:
pri <- jabbaPriors(stk, method = "fishlife")
# or: pri <- jabbaPriors(stk, method = "flbrp")
```

Use this workflow when you want priors grounded in age-structured demography,
especially for data-limited stocks where life-history assumptions drive
productivity expectations.

# Side-by-side comparison

```{r compare}
toDf <- function(x, methodName) {
  muVals <- as.numeric(c(x$pr))
  muNames <- dimnames(x$pr)$params
  names(muVals) <- muNames

  sdVals <- as.numeric(c(x$pr.sd))
  sdNames <- dimnames(x$pr.sd)$params
  names(sdVals) <- sdNames

  data.frame(
    method = methodName,
    param = muNames,
    mean = as.numeric(muVals),
    sd = as.numeric(sdVals[match(muNames, sdNames)])
  )
}

cmp <- rbind(
  toDf(priIces, "ices"),
  toDf(priFishlife, "fishlife"),
  toDf(priFlbrp, "flbrp")
)

cmp
```

# Practical guidance

- Start with `"ices"` when ICES attributes are complete and trusted.
- Use `"fishlife"` when you need a biologically informed fallback.
- Use `"flbrp"` when your workflow is explicitly anchored in the FLBRP/equilibrium fit.
- Compare methods as a sensitivity check before final JABBA runs.

# Next step: fit model

Once priors are selected:

```{r next-step, eval=FALSE}
fit <- runJABBA(stk, method = "ices", quick = TRUE, nc = 3)
```

For full workflow details, see the vignette:

- `vignettes/jabba-workflow.Rmd`
