---
title: "Analytic catch projections (JABBA)"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    code_folding: hide
VignetteBuilder: knitr
vignette: >
  %\VignetteIndexEntry{Analytic catch projections (JABBA)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 4,
  dpi = 120, warning = TRUE, message = FALSE)
```

# Purpose

Same products as the Stock Synthesis vignette, from a JABBA fit:
`projectJabba()` catch forecasts, `kobeMomentGridJabba()` log-moments,
`makeK2SM()`, and `kobeRebuildCurve()`. Intervals use `kobeMomentCI()`
(same lognormal assumption as `lognormalCI()`).

Set `FLREBUILD_JABBADIR` / `FLREBUILD_JABBARUN`. Defaults are the 2026
North Atlantic shortfin mako scenario `1-S`.

```{r setup, warning=FALSE}
if (requireNamespace("pkgload", quietly = TRUE) &&
    file.exists(file.path("..", "DESCRIPTION"))) {
  pkgload::load_all("..", quiet = TRUE, export_all = FALSE)
} else {
  library(FLRebuild)
}
library(ggplot2)

hasKobe  <- requireNamespace("kobe", quietly = TRUE)
hasJabba <- requireNamespace("JABBA", quietly = TRUE)
jabbaDir <- Sys.getenv(
  "FLREBUILD_JABBADIR",
  unset = paste0(
    "P:/rfmo/iccat/2026/SHK/Analysis/final assessment results/",
    "JABBA/final3JABBA/outputCorrected"))
jabbaRun <- Sys.getenv("FLREBUILD_JABBARUN", unset = "1-S")
jabbaExists <- hasJabba && dir.exists(jabbaDir) &&
  file.exists(file.path(jabbaDir, paste0("SMA2026_", jabbaRun, "_jabba.rdata")))

k2Yrs   <- c(2027, 2028, 2029, 2030, 2035, 2040, 2045,
             2050, 2055, 2060, 2065, 2070)
fig2Yrs <- c(2026, 2028, 2030, 2050, 2070)
annYrs  <- 2025:2070
tacsFig <- c(0, 250, 1200, 1883, 2300)
tacs    <- sort(unique(c(tacsFig, seq(0, 2300, by = 100))))
LEVELS  <- c(0.5, 0.8, 0.95)

theme_kobe <- theme_bw(base_size = 10) +
  theme(panel.grid = element_blank(), legend.position = "bottom")
alpha_levels <- scale_alpha_manual(
  values = c("0.5" = 0.35, "0.8" = 0.22, "0.95" = 0.12), name = "Level")
```

```{r load-project, eval=jabbaExists}
remoteFit <- file.path(jabbaDir, paste0("SMA2026_", jabbaRun, "_jabba.rdata"))
localFit  <- file.path(tempdir(), basename(remoteFit))
if (!file.exists(localFit) ||
    !identical(unname(file.info(localFit)$size), unname(file.info(remoteFit)$size)))
  stopifnot(file.copy(remoteFit, localFit, overwrite = TRUE))

fit <- getRuns(setNames(localFit, jabbaRun), source = "jabba")[[jabbaRun]]
termYr <- max(fit$kbtrj$year, na.rm = TRUE)
prj <- projectJabba(fit, catch = tacs, nyears = 2070L - termYr,
                    imp.yr = 3, thin = 10, prjOnly = TRUE)
```

# Trajectories

JABBA `kbtrj` through the terminal year, then `projectJabba()` forecasts.
Ribbons from `kobeMomentCI()` at 50 / 80 / 95%.

```{r fig1-data, eval=jabbaExists}
kbH <- jabbaKbtrj(fit, yearMax = termYr)
histFig <- do.call(rbind, lapply(tacsFig, function(c0) {
  x <- kbH[, c("year", "stock", "harvest")]
  x$catch <- c0
  x
}))
figTrj <- rbind(histFig, prj[prj$catch %in% tacsFig, names(histFig)])
gridFig <- kobeMomentGridJabba(figTrj, years = sort(unique(figTrj$year)))
ssbCI <- transform(kobeMomentCI(gridFig, "B", LEVELS),
                   catch = factor(catch, levels = tacsFig))
fCI   <- transform(kobeMomentCI(gridFig, "F", LEVELS),
                   catch = factor(catch, levels = tacsFig))
ssbMed <- ssbCI[ssbCI$level == 0.5, ]
fMed   <- fCI[fCI$level == 0.5, ]
```

```{r fig1-ssf, eval=jabbaExists, fig.height=3}
ggplot(ssbCI, aes(year, ratio)) +
  geom_ribbon(aes(ymin = ratioLo, ymax = ratioHi, alpha = factor(level)),
              fill = "#1b9e77") +
  geom_line(data = ssbMed, linewidth = 0.7) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40") +
  geom_hline(yintercept = 0.4, linetype = "3313", colour = "red4") +
  geom_vline(xintercept = termYr + 0.5, linetype = "dotted", colour = "grey50") +
  facet_wrap(~catch, nrow = 1, labeller = label_both) +
  labs(x = NULL, y = expression(B / B[MSY])) +
  coord_cartesian(ylim = c(0, 3)) + alpha_levels + theme_kobe
```

```{r fig1-f, eval=jabbaExists, fig.height=3}
ggplot(fCI, aes(year, ratio)) +
  geom_ribbon(aes(ymin = ratioLo, ymax = ratioHi, alpha = factor(level)),
              fill = "#d95f02") +
  geom_line(data = fMed, linewidth = 0.7) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40") +
  geom_vline(xintercept = termYr + 0.5, linetype = "dotted", colour = "grey50") +
  facet_wrap(~catch, nrow = 1, labeller = label_both) +
  labs(x = NULL, y = expression(F / F[MSY])) +
  coord_cartesian(ylim = c(0, 5)) + alpha_levels + theme_kobe
```

# Kobe II strategy matrix

```{r table1, eval=jabbaExists, fig.height=12, fig.width=6, warning=FALSE}
k2 <- makeK2SM(kobeMomentGridJabba(prj, years = k2Yrs), years = k2Yrs)
k2sm <- rbind(
  cbind(k2$long, panel = "P(F <= Fmsy)", p = k2$long$pNoOverfishing),
  cbind(k2$long, panel = "P(B >= Bmsy)", p = k2$long$pNotOverfished),
  cbind(k2$long, panel = "P(green)", p = k2$long$green))
k2sm$panel <- factor(k2sm$panel, unique(k2sm$panel))
k2sm$catch <- factor(k2sm$catch, sort(unique(k2sm$catch)))
k2sm$year  <- factor(k2sm$year, k2Yrs)

ggplot(k2sm, aes(year, catch, fill = p)) +
  geom_tile(colour = "grey85", linewidth = 0.2) +
  geom_text(aes(label = sprintf("%.0f", 100 * p)), size = 2.2, colour = "#00045b") +
  scale_fill_gradientn(
    colours = c(colorRampPalette(c("red4", "red"))(12),
                colorRampPalette(c("yellowgreen", "darkgreen"))(8)),
    limits = c(0, 1), name = "P") +
  scale_y_discrete(limits = rev(levels(k2sm$catch))) +
  facet_wrap(~panel, ncol = 1) +
  labs(x = NULL, y = "Catch (t)") + theme_kobe
```

# Densities

```{r fig2, eval=jabbaExists, fig.height=2.5}
grid2 <- kobeMomentGridJabba(prj[prj$catch %in% tacsFig, ], years = fig2Yrs)
dens_df <- do.call(rbind, lapply(seq_len(nrow(grid2)), function(i) {
  g <- grid2[i, ]
  out <- list()
  if (is.finite(g$sigmaB) && g$sigmaB > 0)
    out[[1]] <- data.frame(
      catch = g$catch, year = g$year, quantity = "B/Bmsy",
      x = exp(seq(log(0.2), log(3), length.out = 200)),
      dens = dlnorm(exp(seq(log(0.2), log(3), length.out = 200)),
                    g$muB, g$sigmaB))
  if (is.finite(g$sigmaF) && g$sigmaF > 0 && is.finite(exp(g$muF)))
    out[[2]] <- data.frame(
      catch = g$catch, year = g$year, quantity = "F/Fmsy",
      x = exp(seq(log(0.05), log(4), length.out = 200)),
      dens = dlnorm(exp(seq(log(0.05), log(4), length.out = 200)),
                    g$muF, g$sigmaF))
  if (length(out)) do.call(rbind, out)
}))
```

```{r fig2-ssf, eval=jabbaExists, fig.height=2.5}
ggplot(subset(dens_df, quantity == "B/Bmsy"),
       aes(x, dens, colour = factor(catch))) +
  geom_line(linewidth = 0.6) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
  facet_wrap(~year, nrow = 1, scales = "free_y") + scale_x_log10() +
  labs(x = expression(B / B[MSY]), y = "Density", colour = "Catch (t)") +
  theme_kobe
```

```{r fig2-f, eval=jabbaExists, fig.height=2.5}
ggplot(subset(dens_df, quantity == "F/Fmsy"),
       aes(x, dens, colour = factor(catch))) +
  geom_line(linewidth = 0.6) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
  facet_wrap(~year, nrow = 1, scales = "free_y") + scale_x_log10() +
  labs(x = expression(F / F[MSY]), y = "Density", colour = "Catch (t)") +
  theme_kobe
```

# Phase plot

```{r kobe-phase-mar, eval=jabbaExists && hasKobe, fig.height=6, fig.width=6, warning=FALSE}
library(kobe)
g2070 <- kobeMomentGridJabba(prj[prj$catch %in% tacsFig, ], years = 2070)
kobe::kobePhaseMar(kobeDraws(g2070, n = 800, seed = 2070),
                   xlab = expression(B / B[MSY]),
                   ylab = expression(F / F[MSY]))
```

```{r kobe-phase-fallback, eval=jabbaExists && !hasKobe}
message("Install kobe (github.com/flr/kobe) for kobePhaseMar.")
```

# Recovery curve

Last year \(P(\mathrm{green})\) stays at 0.50 or 0.60, by catch.

```{r rebuild, eval=jabbaExists, fig.height=3.5, warning=FALSE}
gridAnn <- kobeMomentGridJabba(prj, years = annYrs)
curve <- kobeRebuildCurve(gridAnn, catch = tacs, years = annYrs,
                          target = c(0.50, 0.60))
ggplot(curve, aes(catch, rebuildYear, colour = factor(target))) +
  geom_line() + geom_point(size = 1.4) +
  labs(x = "Catch (t)", y = "Rebuild year", colour = "P(green)") +
  theme_kobe
```

```{r status}
if (!jabbaExists)
  message("Need JABBA + fit at ", jabbaDir, "; projection chunks skipped.")
```
