---
title: "Unified Prior Thread: JABBA, SPiCT, and FLICC"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Unified Prior Thread: JABBA, SPiCT, and FLICC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Purpose

This vignette outlines a single prior-generation thread that can feed:

- JABBA
- SPiCT
- FLICC

The key idea is to first build a canonical prior object from one of several
pathways (`ices`, `fishlife`, `flbrp`, `lhPar`/`lhEql`, Leslie matrix), then
map that object into model-specific prior inputs.

# Why a unified thread

A shared thread helps:

- keep biological assumptions consistent across models
- document provenance of each prior
- run sensitivity analyses in a reproducible way
- avoid hidden differences in model defaults

# Setup

```{r setup}
library(FLRebuild)
library(icesdata)

data(icesdata)
stk <- icesdata[[1]]
```

# Step 1: Build canonical priors

Use one of the available pathways:

- `jabbaPriors(stk, method = "ices")`
- `jabbaPriors(stk, method = "fishlife")`
- `jabbaPriors(stk, method = "flbrp")`
- life-history to FLBRP route (`lhPar`, `lhEql`)
- Leslie matrix route (`FLife::leslie`)

```{r canonical-prior-example}
pri <- FLRebuild::jabbaPriors(stk, method = "ices")
pri$pr
pri$pr.sd
```

Create a canonical table with core quantities. This is a lightweight structure
you can save and pass to adapter functions:

```{r canonical-table}
canonical <- FLRebuild::canonicalPriors(stk, method = "ices")

canonical
```

# Step 2: Optional uncertainty layer (CV and covariance)

You can enrich canonical priors with:

- `cv`: coefficient of variation (`sd / mean`)
- covariance matrix: from `sd` and assumed/estimated correlations

```{r cv-and-covariance}
# Basic independent assumption (diagonal covariance):
covMatIndependent <- FLRebuild::priorCovariance(canonical)

canonical[, c("param", "mean", "sd", "cv")]
covMatIndependent
```

If you have correlation assumptions, build a correlation matrix and convert to
covariance:

```{r covariance-with-correlation, eval = FALSE}
rho <- diag(nrow(canonical))
dimnames(rho) <- list(canonical$param, canonical$param)

# Example: set correlation between two parameters
# rho["r", "shape"] <- 0.3
# rho["shape", "r"] <- 0.3

sdDiag <- diag(canonical$sd)
covMat <- FLRebuild::priorCovariance(canonical, correlation = rho)
```

# Step 3: Add provenance metadata

Track where priors came from and what assumptions were used:

```{r provenance}
priorMeta <- list(
  stockId = names(icesdata)[1],
  pathway = "ices",
  notes = "Baseline run for cross-model comparison",
  generatedOn = as.character(Sys.Date())
)

priorMeta
```

# Step 4: Map to model-specific priors (adapters)

Use package adapters so each model gets priors in the expected shape.

```{r adapters, eval = FALSE}
jabbaPrior <- FLRebuild::asJabbaPriors(canonical)
spictPrior <- FLRebuild::asSpictPriors(canonical, covariance = covMatIndependent)
fliccPrior <- FLRebuild::asFliccPriors(canonical, covariance = covMatIndependent)
```

For a FLICC-specific YPR-driven route, use:

```{r flicc-ypr-priors}
fliccYpr <- FLRebuild::asFliccPriorsYpr(stk)
fliccYpr$reference
head(fliccYpr$curve)
```

# Step 5: Run model entry points from one prior thread

```{r model-entrypoints, eval = FALSE}
fitJabba <- FLRebuild::runJABBA(
  stk,
  method = "ices",
  priors = jabbaPrior,
  quick = TRUE,
  nc = 3
)

# SPiCT (illustrative)
spictPrior <- FLRebuild::asSpictPriors(canonical, covariance = covMatIndependent)
# fitSpict <- fit.spict(data = ..., priors = spictPrior)

# FLICC (illustrative)
fliccPrior <- FLRebuild::asFliccPriors(canonical, covariance = covMatIndependent)
# fitFlicc <- FLRebuild::fitFLICC(stk, ...)
```

# Step 6: Sensitivity over pathways

Build multiple canonical priors and compare fits under each pathway:

```{r sensitivity-template, eval = FALSE}
pathways <- c("ices", "fishlife", "flbrp")

priorsByPathway <- lapply(pathways, function(mth) {
  cc <- FLRebuild::canonicalPriors(stk, method = mth, source = mth)
  transform(cc, pathway = mth)
})
names(priorsByPathway) <- pathways

do.call(rbind, priorsByPathway)
```

# Recommended project pattern

- Keep one canonical prior object per stock and pathway
- Version that object with metadata and date
- Store CV and covariance assumptions explicitly
- Use explicit adapters for each model
- Compare pathway-induced differences before final advice
